---
title: "FWO postdoctoral research proposal"
subtitle: "Information on Work Package 1"
date: 'November 28, 2024' #"Latest update - `r format(Sys.time(), '%d %B %Y')`"
# author:
#   - Joachim Waterschoot^[Department of Developmental, Personality and Social Psychology, Ghent University, joachim.waterschoot@ugent.be]
output:
  html_document:
    toc: false
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: yeti
    highlight: textmate
    number_sections: false
    # includes:
    #   after_body: footnote.html
  # pdf_document:
  #   toc: yes
    # latex_engine: xelatex
    # number_sections: false
always_allow_html: true
mainfont: Aptos
classoption:
- twocolumn
---

<!-- \fontsize{9}{10} -->
\selectfont

<style type="text/css">
<!-- body{ -->
<!--   font-size: 10pt; -->
<!-- } -->
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
div.orange { background-color:#ffa366; border-radius: 5px; padding: 20px;}
div.yellow { background-color:#09bb9f; color: white; border-radius: 5px; padding: 20px;}
div.grey { background-color:#ebebeb; border-radius: 5px; padding: 20px;}

  .plotly {
    width: 100% !important;
    height: auto !important;
  }

</style>

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

<br> 

# Overview

<div class = "col-md-6">
<div class = "blue">

As part of Work Package 1 in the FWO Postdoctoral (junior), we have confirmation to collaborate of

+ **57** countries

+ **110** collaborators

+ proportionally more democratic than full list of countries (see figure right)

On this webpage, maps can checked of these countries with their *Democracy Index* scores and a list of the collaborators, including their signed form. 

</div>
</div>
<div class = "col-md-6">



<center> 
```{r echo=FALSE, warning=FALSE,message = FALSE, out.width="100%"}
library(dplyr)
library(ggplot2)


Rnew <- readxl::read_excel('/Users/joachimwaterschoot/Library/CloudStorage/OneDrive-UGent/RESEARCH/FWO/Engagements/Overview.xlsx')
Rnew <- as.data.frame(Rnew)
Rnew <- Rnew[-c(1),c(1:14)]
names(Rnew) <- Rnew[1,]
Rnew <- Rnew[-1,]

library(flextable)

Rnew <- Rnew[,c('Name','Institution','Country','Flag',"Continent","Democracy index","Type","Hyperlink")]
Rnew$`Democracy index` <- round(as.numeric(Rnew$`Democracy index`),2)
row.names(Rnew) <- NULL
Rnew <- Rnew[order(Rnew$Country,decreasing=FALSE),]
Rnew <- Rnew[1:(which(Rnew$Name=='end')-1),]

Rnew <- unique(Rnew[,c('Country',"Type")])
Rnew <- as.data.frame(Rnew)
colnames(Rnew)[which(colnames(Rnew)=='Type')] <- 'category'
colnames(Rnew)[which(colnames(Rnew)=='Country')] <- 'country'
your_data <- Rnew
colnames(your_data) <- c('country','category')
your_data$category <- as.factor(your_data$category)
your_data$category <- factor(your_data$category, levels=c("Autocratic regime","Hybrid regime","Flawed democracy","Full democracy" ))

# Full dataset of all countries and categories
Rnew <- readxl::read_excel('/Users/joachimwaterschoot/Library/CloudStorage/OneDrive-UGent/RESEARCH/FWO/Engagements/Overview.xlsx',sheet = 'DemInd2023')
Rnew <- as.data.frame(Rnew)
full_data <- na.omit(Rnew[,c(1,4)])
colnames(full_data) <- c('country','category')
full_data$category <- as.factor(full_data$category)
full_data$category <- factor(full_data$category, levels=c("Autocratic regime","Hybrid regime","Flawed democracy","Full democracy" ))

# Calculate proportions for each category in your dataset
your_data_summary <- your_data %>%
  group_by(category) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count),
         dataset = "Work Package 1")

# Calculate proportions for each category in the full dataset
full_data_summary <- full_data %>%
  group_by(category) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count),
         dataset = "Full list")

# Combine the summaries
comparison_data <- bind_rows(your_data_summary, full_data_summary)

# Plot comparison as a bar chart
ggplot(comparison_data, aes(x = category, y = proportion, fill = dataset)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Comparison of democracy categories: collaborating countries vs full List",
    x = "Democracy category",
    y = "Proportion of countries",
    fill = "Dataset"
  ) +
  theme_minimal()

```
</center>

</div>

<br> 

# Maps

<span style="color: blue; cursor: help;" title="Hover over the map to see details about each country's governance type!">
Tip: You can zoom in and hover on the map to see more details. Double click to reset
</span>

```{r echo=FALSE, warning=FALSE,message = FALSE}

library(plotly)
library(dplyr)



Rnew <- readxl::read_excel('/Users/joachimwaterschoot/Library/CloudStorage/OneDrive-UGent/RESEARCH/FWO/Engagements/Overview.xlsx')
Rnew <- as.data.frame(Rnew)
Rnew <- Rnew[-c(1),c(1:14)]
names(Rnew) <- Rnew[1,]
Rnew <- Rnew[-1,]

library(flextable)

Rnew <- Rnew[,c('Name','Institution','Country','Flag',"Continent","Democracy index","Type","Hyperlink")]
Rnew$`Democracy index` <- round(as.numeric(Rnew$`Democracy index`),2)
row.names(Rnew) <- NULL
Rnew <- Rnew[order(Rnew$Country,decreasing=FALSE),]
Rnew <- Rnew[1:(which(Rnew$Name=='end')-1),]

Rnew <- unique(Rnew[,c('Country',"Democracy index")])
Rnew <- as.data.frame(Rnew)
colnames(Rnew)[which(colnames(Rnew)=='Democracy index')] <- 'democracy_index'
colnames(Rnew)[which(colnames(Rnew)=='Country')] <- 'country'
data <- Rnew

# Custom colorscale
custom_colorscale <- list(
  c(0, "#f0f9e8"), # Light green for lowest values
  c(1, "#254b8c")  # Dark blue for highest values
)

# Create the plot
plot <- plot_ly(
  data,
  type = 'choropleth',
  locations = ~country,
  locationmode = 'country names',
  z = ~democracy_index,
  colorscale = custom_colorscale,
  text = ~paste(country,"<br>",'Democracy Index (2023) =', democracy_index, sep=""), # Custom hover text
  hoverinfo = "text", # Show only custom text in hover
  colorbar = list(
    title = list(
      text = "Democracy Index (2023)", # Legend title
      font = list(family = "Arial", size = 14)
    ),
    x = 0.5, # Center legend horizontally
    xanchor = "center", # Align legend by center horizontally
    y = 1.15, # Position above the map
    orientation = "h", # Horizontal legend
    thickness = 10
  )
) %>%
  layout(
    geo = list(
         projection = list(type = "natural earth", scale = 1), # Adjust scale for a larger map
      showlakes = TRUE,
      lakecolor = 'rgb(173, 216, 230)'
      #center = list(lon = 0, lat = 20) # Center the map
    ),
    width = NULL, # Allow the plot to be responsive
    height = NULL # Allow the plot to adapt its height
  )

plot
```

<br>

<center> <b>Collaborating countries by regime type (based on Democracy Index, 2023) </b></center>

```{r echo=FALSE, warning=FALSE,message = FALSE}

Rnew <- readxl::read_excel('/Users/joachimwaterschoot/Library/CloudStorage/OneDrive-UGent/RESEARCH/FWO/Engagements/Overview.xlsx')
Rnew <- as.data.frame(Rnew)
Rnew <- Rnew[-c(1),c(1:14)]
names(Rnew) <- Rnew[1,]
Rnew <- Rnew[-1,]

library(flextable)

Rnew <- Rnew[,c('Name','Institution','Country','Flag',"Continent","Democracy index","Type","Hyperlink")]
Rnew$`Democracy index` <- round(as.numeric(Rnew$`Democracy index`),2)
row.names(Rnew) <- NULL
Rnew <- Rnew[order(Rnew$Country,decreasing=FALSE),]
Rnew <- Rnew[1:(which(Rnew$Name=='end')-1),]

Rnew <- unique(Rnew[,c('Country',"Type")])
Rnew <- as.data.frame(Rnew)
colnames(Rnew)[which(colnames(Rnew)=='Type')] <- 'category'
colnames(Rnew)[which(colnames(Rnew)=='Country')] <- 'country'
data <- Rnew

# Assign unique colors for each category
category_colors <- c("Autocratic regime" = "#d73027", 
                     "Hybrid regime" = "#fc8d59", 
                     "Flawed democracy" = "#fee08b", 
                     "Full democracy" = "#1a9850")

# Create a categorical map
plot <- plot_ly(
  data,
  type = 'choropleth',
  locations = ~country,
  locationmode = 'country names',
  z = ~as.numeric(factor(category, levels = names(category_colors))), # Map categories to numeric
  colorscale = list(
    list(0, category_colors["Autocratic regime"]),
    list(0.33, category_colors["Hybrid regime"]),
    list(0.66, category_colors["Flawed democracy"]),
    list(1, category_colors["Full democracy"])
  ),
  text = ~paste('Country:', country, '<br>Category:', category), # Custom hover text
  hoverinfo = "text", # Show only custom hover text
  colorbar = list(
    title = list(
      text = "Governance Category",
      font = list(family = "Arial", size = 14)
    ),
    tickvals = c(1, 2, 3, 4), # Numeric values for categories
    ticktext = names(category_colors), # Corresponding category names
    thickness = 15 # Adjust thickness of the colorbar
  )
) %>%
  layout(
    
    geo = list(
      projection = list(type = "natural earth", scale = 1),
      showlakes = FALSE
    )
  )

plot
```

# List of collaborators

<span style="color: blue; cursor: help;" title="Click on the 'Form' hyperlink to see the signed collaboration form">
Click on the 'Form' hyperlink to see the signed collaboration form
</span>

```{r echo=FALSE, warning=FALSE,message = FALSE}
library(ftExtra)
library(knitr)
library(kableExtra)
library(lubridate)

Rnew <- readxl::read_excel('/Users/joachimwaterschoot/Library/CloudStorage/OneDrive-UGent/RESEARCH/FWO/Engagements/Overview.xlsx')
Rnew <- as.data.frame(Rnew)
Rnew <- Rnew[-c(1),c(1:14)]
names(Rnew) <- Rnew[1,]
Rnew <- Rnew[-1,]

library(flextable)

Rnew <- Rnew[,c('Name','Institution','Country','Flag',"Continent","Democracy index","Type","Hyperlink")]
Rnew$`Democracy index` <- round(as.numeric(Rnew$`Democracy index`),2)
colnames(Rnew)[which(colnames(Rnew)=='Democracy index')] <- 'Democracy index (2023)'
row.names(Rnew) <- NULL
Rnew <- Rnew[order(Rnew$Country,decreasing=FALSE),]
Rnew <- Rnew[1:(which(Rnew$Name=='end')-1),]
df <- Rnew

library(DT)
datatable(df,
          escape=FALSE,
        #  caption = 'Overview of reports',
         # filter = 'top',
          rownames = FALSE,
          extensions = c('ColReorder',
                         'KeyTable','Responsive'),
          options = list(
           # order = list(list(2, 'desc')),
            pageLength = 70,
            autoWidth = TRUE,
            dom = 'Bfrtip',
           # buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
            colReorder = TRUE,
            keys = TRUE
            )
          )  %>%
  formatStyle(columns = names(df), fontSize = '15px') %>%
  # Format the 'Value' column to display integers without separators
  formatStyle(columns = "Year", value = formatC(0, format = "d", big.mark = "", decimal.mark = "."))



```

