---
title: "Correlations"
description: |
  A theoretical and practical guide for K-Means clustering and Latent Profile Analysis in R
author:
  - name: Joachim Waterschoot
  - orcid_id: 0000-0003-0845-9310
date: 06-09-2021
categories:
  - clustering
  - analyses
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
    toc_float: true
    highlight: haddock
    highlight_downlit: true
    code_folding: true
---

<style type="text/css">
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
div.orange { background-color:#ffa366; border-radius: 5px; padding: 20px;}
div.yellow { background-color:#09bb9f; color: white; border-radius: 5px; padding: 20px;}
</style>


<meta name="decsription" content="Find useful online applications to make correlation tables, moderation or interaction plots and many more." />

<!-- Global site tag (gtag.js) - Google Analytics -->
<script async src="https://www.googletagmanager.com/gtag/js?id=UA-120279443-3"></script>
<script>
  window.dataLayer = window.dataLayer || [];
  function gtag(){dataLayer.push(arguments);}
  gtag('js', new Date());

  gtag('config', 'UA-120279443-3');
</script>

<script type="text/javascript">

    $(document).ready(function() {
        document.title = 'Rcoding';
    });

</script>

<style>
.main-container {
  background-color:#d1d9e0;
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>




# Correlations

The following code provides a Pearson correlation table with descriptives. 

```{r,echo=TRUE,message=FALSE,warning=FALSE}

packages_folder <- "/Users/joachimwaterschoot/Downloads/R Packages"
  
packages<- c('foreign','lavaan','tidyverse','cluster',
     'factoextra','dendextend','ape','tidyr',
     'magrittr','clValid',
     'ggfortify','clustree','FactoMineR',
     'corrplot','ggiraphExtra','GGally',
     'knitr','kableExtra','plyr','haven','apaTables',
     'qgraph','dplyr','ComplexHeatmap','dendextend',
     'circlize','RColorBrewer','ggpubr','rstatix',
     'car','broom','hausekeep',"writexl",
     'Publish','data.table','janitor','psych',
     'sjPlot', 'Hmisc','xtable','survey',
     'srvyr','questionr','ggplot2','ggfortify',
     'lme4','lmerTest','foreign','VIM',"mice",'readxl')

```

```{r, echo=FALSE,results='hide',message=FALSE,warning=FALSE}
lapply(packages, require, character.only = TRUE)
```
```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(apaTables)
library(dplyr)
data  <- foreign::read.spss("(2)Data Joachim.sav", 
                  use.value.labels = FALSE, 
                  to.data.frame=TRUE)
data$dummy1_group <- as.numeric(data$Condition)
data$outcome <- data$Pleasure_Interest
data$moderator1 <- data$Indecisiveness
data$moderator2 <- data$Open
data$gender <- data$Geslacht
data$age <- data$Leeftijd
```             

```{r, echo=TRUE,message=FALSE,warning=FALSE}

# select the variables to correlate
correlation_dataset <- data[, c(

  "Extra","Agree","Con","Neur","Open"
  
  )]

# make sure the variables are all numeric
correlation_dataset <- sapply(correlation_dataset,as.numeric)

# make correlation table with descriptives
corrtable <- apa.cor.table(correlation_dataset, table.number=1)
corrtablee <- as.table(corrtable$table.body)
rangedel <- seq(4,dim(corrtablee)[1],3)+2
rangedel <- c(1,3,rangedel[-length(rangedel)])
corrtablee<- corrtablee[c(rangedel),]

kable(corrtablee, "html") %>%
  kable_paper(full_width = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed", "responsive"),
                fixed_thead = T,font_size = 12) %>%
  kable_classic_2()%>%
  column_spec(3, background = "lightgrey")%>%
  column_spec(4, background = "lightgrey")
```

## Subset of correlation table

You can select only a specific number of columns of the correlation table. Beware that the first 3 columns do refer to the list of variables, the mean and the standard deviation. Here, the first three columns and the first column of the correlation matrix is selected.

```{r, echo=TRUE,message=FALSE,warning=FALSE}

corrtablee <- corrtablee[,1:4]
kable(corrtablee, "html") %>%
  kable_paper(full_width = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed", "responsive"),
                fixed_thead = T,font_size = 12) %>%
  kable_classic_2()%>%
  column_spec(3, background = "lightgrey")%>%
  column_spec(4, background = "lightgrey")
```

## Save it as html

You can save the correlation table as an html file, so you can copy it easily into your manuscript. 

```{r, echo=TRUE,message=FALSE,warning=FALSE}
kable(corrtablee, "html") %>%
  kable_paper(full_width = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed", "responsive"),
                fixed_thead = T,font_size = 12) %>%
  kable_classic_2()%>%
  column_spec(3, background = "lightgrey")%>%
  column_spec(4, background = "lightgrey")%>%
  save_kable("correlation_table.html")
```

<br>




***

&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="www.ugent.be/epg/coronastudie">Joachim Waterschoot</a></p>
<p style="text-align: center;"><span style="color: #808080;"><em>Joachim.Waterschoot@ugent.be</em></span></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- Add font awesome icons -->
<p style="text-align: center;">
    <a href="https://twitter.com/watjoa" class="fa fa-twitter"></a>
    <a href="https://www.linkedin.com/in/watjoa/" class="fa fa-linkedin"></a>
    <a href="https://github.com/watjoa/" class="fa fa-github"></a>
</p>

&nbsp;









